---
title: "e_COMMERCE"
author: "KANNE TAMIBE KOCHIAKE"
date: "2025-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Meta data
Dataset Description: E-commerce Customer Behavior

Overview:
This dataset provides a comprehensive view of customer behavior within an e-commerce platform. Each entry in the dataset corresponds to a unique customer, offering a detailed breakdown of their interactions and transactions. The information is crafted to facilitate a nuanced analysis of customer preferences, engagement patterns, and satisfaction levels, aiding businesses in making data-driven decisions to enhance the customer experience.

Columns:
 - Customer ID (Type: Numeric):
    Description: A unique identifier assigned to each customer, ensuring distinction across the dataset.

- Gender: Type: Categorical (Male, Female)
    Description: Specifies the gender of the customer, allowing for gender-based analytics.

- Age: Type: Numeric
    Description: Represents the age of the customer, enabling age-group-specific insights.

 - City: Type: Categorical (City names)
    Description: Indicates the city of residence for each customer, providing geographic insights.

- Membership Type: Type: Categorical (Gold, Silver, Bronze)
    Description: Identifies the type of membership held by the customer, influencing perks and benefits.

- Total Spend: Type: Numeric
  Description: Records the total monetary expenditure by the customer on the e-commerce platform.

- Items Purchased: Type: Numeric
  Description: Quantifies the total number of items purchased by the customer.

- Average Rating: Type: Numeric (0 to 5, with decimals)
        Description: Represents the average rating given by the customer for purchased items, gauging satisfaction.

- Discount Applied: Type: Boolean (True, False)
    Description: Indicates whether a discount was applied to the customer's purchase, influencing buying behavior.

- Days Since Last Purchase: Type: Numeric
    Description: Reflects the number of days elapsed since the customer's most recent purchase, aiding in retention analysis.

- Satisfaction Level: Type: Categorical (Satisfied, Neutral, Unsatisfied)
    Description: Captures the overall satisfaction level of the customer, providing a subjective measure of their experience.
    

# LOad necessaries libraries

```{r}
library(dplyr)
library(ggplot2)
library(scales)  
```

    
# LOad the data set

```{r}
df = read.csv("E-commerce_Customer_Behavior.csv")
str(df)
summary(df)
```

This data set counts 11 characteristics with 350 custmers. These characteristics are categoricals (Gender,City ,Membership.Type,  Satisfaction.Level ), logique variable ( Discount.Applied) and numericals (Age, Total.spend, Items.purchased,  Average.Rating and Days.Since.Last.Purchase).

Let's transform all catagorical variable as factor.

```{r}
df$Gender = as.factor(df$Gender)
df$City = as.factor(df$City)
df$Membership.Type = as.factor(df$Membership.Type)
df$Satisfaction.Level = as.factor(df$Satisfaction.Level)
```

## EXploratory of data

### Univariate

```{r}
summary(df)
duplicated(df)
```

_Gender_

```{r}
df_counts <- df %>%
  group_by(Gender) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

#barplot
ggplot(df_counts, aes(x = Gender, y = count, fill = Gender)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5) +
  labs(title = "Barplot of City", x = "Gender", y = "Observations") +
  theme_minimal()

```
We have as many women as men in our dataset.

_City_

```{r}

df_counts <- df %>%
  group_by(City) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

#barplot
ggplot(df_counts, aes(x = City, y = count, fill = City)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5) +
  labs(title = "Barplot of City", x = "CIty", y = "Observations") +
  theme_minimal()

```
These customers are from 06 city that are Chicago, Houston, LOs angles, Miami, New york and San Francisco. We have more people who are from New york (16.9%) and LOs ANgeles (16.9%) than anothers city.

_Membership.Type_
```{r}
df_counts <- df %>%
  group_by(Membership.Type) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

#barplot
ggplot(df_counts, aes(x = Membership.Type, y = count, fill = Membership.Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5) +
  labs(title = "Barplot of Membership.Type", x = "Membership.Type", y = "Observations") +
  theme_minimal()
```

We have 03 types's membership of the customer, and their numbers are almost the same (33.1% for Bronze, 33.4% for Gold and 33.4% for Silver)



_Satisfaction.Level_
```{r}
#summary(df$Satisfaction.Level)
df_counts <- df %>%
  group_by(Satisfaction.Level) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

#barplot
ggplot(df_counts, aes(x = Satisfaction.Level, y = count, fill = Satisfaction.Level)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5) +
  labs(title = "Barplot of Satisfaction.Level", x = "Satisfaction.Level", y = "Observations") +
  theme_minimal()
```
We have 03 levels of satisfactions: Neutral, Satisfied and Unsatified. Also, there are some customers who didn't give their satisfaction's level. There are more people who are satified (35.7%)  than the other level (30.6% for Neutral, 33.1% for Unsatified, and 0.6% for those who didn't give their levels.).


_Age_

```{r}
summary(df$Age)
ggplot(df, aes(x=Age)) + geom_boxplot() + labs(title = "Barplot of Age", x="Age", y="Count")
```
THe customer's age vary between 26 and 43 years old. And the clientele is mostly made up of young to middle-aged adults, mainly in the second age bracket (30-40 years).

# TOtal spend

```{r}
summary(df$Total.Spend)
ggplot(df, aes(x=Total.Spend)) + geom_boxplot() + labs(title="Barplot of Toatl.Spend", x="TOtal.Spend", y="Count")
```
There is a wide dispersion of expenditure between customers. The spend vary between $410,8 and $1 520, AS there is a diffferemce between the mean and the median it means that there are some customers who have the big spend. Also 25% des clients dépensent plus de  1 160,6.
 
 
```{r}
summary(df$Items.Purchased)

ggplot(df, aes(x=Items.Purchased)) + geom_boxplot()+ labs(title = "Boxplot of Items.Purchased", x="Itims.Purchased", y = "Count")
summary(df)
```
 










### Bivariate

```{r}
library(corrplot)
    corr_matrix<-cor(df[, c(3,6,8,10)])
    corrplot(corr_matrix,method = "circle",type = "upper", tl.col = "black",  addCoef.col = "black")
```
We note that there is a strong positive dependency (0.94) between Average.Rating and Total.Spend. This means that one increases with the growth of the other.
Average.Rating also decreases with increasing age (cor = -0.72). 
The same applies to depnese, which decreases with increasing age (-0.68).
The greater the Days.since.Last.Purchase., the lower the expenses ( cor=-0.54).

### Analsis of expenditure

```{r}
# Moyenne des dépenses par type de membership
df %>% 
  group_by(Membership.Type) %>% 
  summarise(Mean_Spend = mean(Total.Spend), Count = n()) %>% 
  mutate(Percentage = percent(Count / sum(Count))) %>% 
  ggplot(aes(x=Membership.Type, y=Mean_Spend, fill=Membership.Type)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Percentage), vjust=-0.5, color="black", size=5) +
  scale_fill_manual(values=c("Bronze"="#CD7F32", "Gold"="#FFD700", "Silver"="#C0C0C0")) +
  ggtitle("Dépenses moyennes par type de membership")

# Depense moyenne par ville
df %>% 
  group_by(City) %>% 
  summarise(Mean_Spend = mean(Total.Spend), Count = n()) %>% 
  mutate(Percentage = percent(Count / sum(Count))) %>% 
  ggplot(aes(x=City, y=Mean_Spend, fill=City)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Percentage), vjust=-0.5, color="black", size=5) +
  scale_fill_manual(values=c("Chicago"="#CD7F32", "Houston"= "yellow","Los Angeles"= "green", "Miami"="#C0C0C0", "New York"="steelblue", "San Francisco"="#FFD700")) +
  ggtitle("Dépenses moyennes par ville")

# Boxplot des dépenses en fonction du genre
ggplot(df, aes(x=Gender, y=Total.Spend, fill=Gender)) + 
  geom_boxplot() + 
  scale_fill_manual(values=c("Female"="grey", "Male"="#1E90FF")) +
  ggtitle("Dépenses totales selon le genre")


#spending by age
df$G_age = ifelse(df$Age<= 30 & df$Age> 26,"[26-30years]", "[31-end]")
ggplot(df, aes(x=G_age, y=Total.Spend, fill=G_age)) + 
  geom_boxplot() + 
  scale_fill_manual(values=c("[26-30years]"="grey", "[31-end]"="#1E90FF")) +
  ggtitle("Total spend over Age")

```

**Interretation**
 - In average Gold (more than $1250 ) menbership spend more than silver (almost $750) and Bronze (less than $500). The Gold is our potential clients that spend a lot in our shop.
 
 - Men tend to spend more than women.
 - San Francisco ($1459,772) is the biggest spender, followed by New York (average $1165,036) and Houston (less than $500).
 
 - The average spending of those aged between 26 and 30 is very different from that of those aged between 31 and 43.

```{r}

# Relation entre la ville et le nombre d'articles achetés
df %>% 
  group_by(City) %>% 
  summarise(Mean_Spend = mean(Items.Purchased), Count = n()) %>% 
  mutate(Percentage = percent(Count / sum(Count))) %>% 
  ggplot(aes(x=City, y=Mean_Spend, fill=City)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Percentage), vjust=-0.5, color="black", size=5) +
  scale_fill_manual(values=c("Chicago"="#CD7F32", "Houston"= "yellow","Los Angeles"= "green", "Miami"="#C0C0C0", "New York"="steelblue", "San Francisco"="#FFD700")) +
  ggtitle("Items.Purchased par ville")

# Relation entre la satisfaction et le nombre d'articles achetés
df %>%
  group_by(Satisfaction.Level) %>%
  summarise(Mean_Items = mean(Items.Purchased), Count = n()) %>%
  mutate(Percentage = percent(Count / sum(Count))) %>%
  ggplot(aes(x=Satisfaction.Level, y=Mean_Items, fill=Satisfaction.Level)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Percentage), vjust=-0.5, color="black", size=5) +
  scale_fill_manual(values=c("Neutral"="#F0E68C", "Satisfied"="#32CD32", "Unsatisfied"="steelblue")) +
  ggtitle("Nombre d'articles achetés selon la satisfaction")

# Items.Purchased par type de membership
df %>% 
  group_by(Membership.Type) %>% 
  summarise(Mean_Spend = mean(Items.Purchased), Count = n()) %>% 
  mutate(Percentage = percent(Count / sum(Count))) %>% 
  ggplot(aes(x=Membership.Type, y=Mean_Spend, fill=Membership.Type)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=Percentage), vjust=-0.5, color="black", size=5) +
  scale_fill_manual(values=c("Bronze"="#CD7F32", "Gold"="#FFD700", "Silver"="#C0C0C0")) +
  ggtitle("Items.Purchased par type de membership")


# Boxplot Items.Purchased en fonction du genre
ggplot(df, aes(x=Gender, y=Items.Purchased, fill=Gender)) + 
  geom_boxplot() + 
  scale_fill_manual(values=c("Female"="grey", "Male"="#1E90FF")) +
  ggtitle("Items.Purchased selon le genre")


#Items.Purchased by age
#df$G_age = ifelse(df$Age<= 30 & df$Age> 26,"[26-30years]", "[31-end]")
ggplot(df, aes(x=G_age, y=Items.Purchased, fill=G_age)) + 
  geom_boxplot() + 
  scale_fill_manual(values=c("[26-30years]"="grey", "[31-end]"="#1E90FF")) +
  ggtitle("Items.Purchased over Age")
```

  
- Satisfied customers are those who buy the most products on average. There are also some people who buy on average more than 10 item but they aren't satified.
 
 
## Forcasting
 
```{r}
df2 = df[, -c(1,8,9,11,5,12)]
model = lm(Total.Spend ~ ., data=df2)
summary(model)

```
```{r}
df2 = df2[, -1]
model2 = lm(Total.Spend ~ ., data=df2)
summary(model2)
```

 **Interpretation**:
- For each additional year of age, total expenditure increases by an average of 1.74 units (e.g. dollars), all else being equal.
- Customers in San Francisco and New York spend significantly more than those in other cities.
- For each additional item purchased, total expenditure increases by an average of 25.33 units.
- For each additional day since the last purchase, total expenditure decreases by an average of 1.28 units.

**Implications for the company**

 _Priority targets:_
 Customers in San Francisco and New York spend significantly more. The company could focus its marketing efforts on these regions.
Older customers and those who buy more items are also priority targets.

_Loyalty:_
 Customers who return more frequently (fewer days since last purchase) spend more. Loyalty programs could encourage more frequent purchases.

_Gender:__
Gender has no significant impact on spending. Marketing strategies should therefore not be differentiated according to gender.

 
 
 
```{r}
anova(model, model2)
```


